---
title: "Contingency Analyses"
author: "Argyris Stringaris"
date: "31/01/2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Think about how best to test the alignment hypothesis

# Import data
```{r}
library(tidyverse)
library(reshape2)
Experiment3 <- read.csv("~/Downloads/Experiment3_cleaned.csv")
# Note: these are the cleaned data that Song sent me on 30th Jan 2022 with explanations in Markdown
dim(Experiment3)

```

# Number of subjects
```{r}
length(unique(Experiment3$subject))
# Note: Song, the previous dataset had 80 subjects. We must be able to account why we excluded any 
# subject from our original MTurk collection
```

# Number of stimuli       
```{r}
length(unique(Experiment3$type2))
#Note: Song, the previous dataset had over 30 stimuli. We need to be clear how many stimuli the subjects really saw. Where there 3 test ones? We should include and label them. What were the other ones? 
```


# Create new contingency and choice variables with labels
```{r}
Experiment3$contingency <- ifelse(Experiment3$contingency2 == 1, "weapon", "non_weapon")
Experiment3$choice <- ifelse(Experiment3$choice2 == 1, "fight", "escape")
#head(Experiment3[, c("contingency2", "contingency", "choice2", "choice")], -10)
```


# Categorical presentation of choice by contingency
```{r}
cross_tab_choice_contingency <-table(Experiment3$contingency,Experiment3$choice)

prop_choice_contingency <- data.frame(prop.table(cross_tab_choice_contingency,1)*100)

colnames(prop_choice_contingency) <- c("contingency", "choice", "percentage")

prop_choice_contingency$contingency <- factor(prop_choice_contingency$contingency)

prop_choice_contingency_chisq <- chisq.test(Experiment3$contingency,Experiment3$choice)
prop_choice_contingency_chisq_stats <- round(with(prop_choice_contingency_chisq, c(statistic, parameter, p.value)), 3)

prop_choice_contingency$counts <- c(cross_tab_choice_contingency[1,1], cross_tab_choice_contingency[2,1],cross_tab_choice_contingency[1,2],    cross_tab_choice_contingency[2,2])

prop_choice_contingency
```


# Graph categorical presentation of choice and contingency
```{r}
prop_choice_contingency %>% 
  ggplot(aes(x = contingency, y = percentage, fill = choice)) +
  geom_bar(stat="identity", position=position_dodge(.93)) +
  ggtitle("choice by contingency") +
  ylab("percentage chosen") +
annotate("text", x = 1, y = 75, label = paste0("χ2 = ", prop_choice_contingency_chisq_stats[1], ", df = ", prop_choice_contingency_chisq_stats[2], ", p = ", prop_choice_contingency_chisq_stats[3] ))
```

# Check distribution of anger
```{r}
hist(Experiment3$anger_rating)

```



# Look at average levels of anger by contingency
```{r}
levels_contingency <- levels(as.factor(Experiment3$contingency))
avg_anger_rating <- 0
sd_anger_rating <- 0
for (i in seq(levels_contingency)){
  avg_anger_rating[i] <-  mean(Experiment3[Experiment3$contingency==levels_contingency[i],]$anger_rating)
  sd_anger_rating[i] <-  sd(Experiment3[Experiment3$contingency==levels_contingency[i],]$anger_rating)
avg_anger_ratings_by_contingency <- data.frame(cbind(levels_contingency, avg_anger_rating, sd_anger_rating ))
}
avg_anger_ratings_by_contingency
# Note the high sd 
```


# Plot the anger mean differences between contingencies  and t-test results
```{r}
test_mean_anger <- t.test(Experiment3$anger_rating ~ Experiment3$contingency, paired = TRUE)

Experiment3 %>% 
  ggplot(aes(x = contingency, y = anger_rating)) +
  geom_boxplot() + 
  annotate("text", x = 1.5, y = 2.5, label = paste0("t = ", round(test_mean_anger$statistic[1],2), ", df = ,", test_mean_anger$parameter, "\n", "p = ", test_mean_anger$p.value ))
# these are very small differences
```


# Plot the anger mean differences between contingencies broken down by choice
```{r}
Experiment3 %>% 
  ggplot(aes(x = contingency, y = anger_rating)) +
  geom_boxplot() +
  facet_grid(~choice)
# and there is little evidence of a robust interaction by choice
```


# Get positive slopes for each subject
```{r}

slopes <- matrix(NA, nrow =  length(unique(Experiment3$subject)), ncol = length(unique(Experiment3$type2)))

for (j in seq(unique(Experiment3$subject))){
for (i in seq(unique(Experiment3$type2))) {
  
 # if(is.na(test[test$contingency=="weapon" & test$type2 == i ,]$anger_rating)==FALSE & #is.na(test[test$contingency=="non_weapon" & test$type2 == i ,]$anger_rating)==FALSE)
  
slopes[j,i] <- Experiment3[Experiment3$contingency=="weapon" & Experiment3$type2 == i & Experiment3$subject == j ,]$anger_rating - Experiment3[Experiment3$contingency=="non_weapon" & Experiment3$type2 == i & Experiment3$subject == j ,]$anger_rating

anger_slopes_df <- data.frame(cbind(subect = unique(Experiment3$subject), slopes))
 colnames(anger_slopes_df) <- c("subject", unique(Experiment3$type2)) 

}
}
dim(anger_slopes_df )

```



# What is the proportion of positive slopes per subject and across subjects?
```{r}
proportion_pos_slopes_per_subject <- 0
for (i in seq(nrow(anger_slopes_df))){
  
  proportion_pos_slopes_per_subject[i] <- 
    
    sum(anger_slopes_df[i,c(2:28)]>0)/length(anger_slopes_df[i,c(2:28)])

}

proportion_pos_slopes_per_subject  

average_positive_slopes_across_subjects <- mean(proportion_pos_slopes_per_subject)
paste0 ("average_positive_slopes_across_subjects = ", print(average_positive_slopes_across_subjects))

# Note: a very small preponderance of positive slopes
```



# step to get a dataframe that denotes which trials amongst the ones where there is a positive contingency (i.e. weapon) had which CHOICE (fight or escape)
```{r}
slopes_weapon <- matrix(NA, nrow =  length(unique(Experiment3$subject)), ncol = length(unique(Experiment3$type2)))



for (j in seq(unique(Experiment3$subject))){
for (i in seq(unique(Experiment3$type2))) {
  
 # if(is.na(test[test$contingency=="weapon" & test$type2 == i ,]$anger_rating)==FALSE & #is.na(test[test$contingency=="non_weapon" & test$type2 == i ,]$anger_rating)==FALSE)
  
slopes_weapon[j,i] <- Experiment3[Experiment3$contingency=="weapon" & Experiment3$type2 == i & Experiment3$subject == j ,]$choice

#anger_slopes_df <- data.frame(cbind(subect = unique(Experiment3$subject), slopes))
 #colnames(anger_slopes_df) <- c("subject", unique(Experiment3$type2)) 

}
}

dim(slopes_weapon)
```



```{r}
nrow(anger_slopes_df[slopes_weapon == "fight", ])
nrow(anger_slopes_df[slopes_weapon == "escape", ])
#these make sense
```

```{r}
positive_slopes_when_weapon_and_fight <- anger_slopes_df[slopes_weapon == "fight", ]
positive_slopes_when_weapon_and_esccape <- anger_slopes_df[slopes_weapon == "escape", ]
```



# positive slopes when presented with a weapon and choosing to fight
```{r}
proportion_pos_slopes_per_subject_fight <- 0
for (i in seq(nrow(positive_slopes_when_weapon_and_fight))){
  
  proportion_pos_slopes_per_subject_fight[i] <- 
    
    sum(positive_slopes_when_weapon_and_fight[i,c(2:28)]>0)/length(positive_slopes_when_weapon_and_fight[i,c(2:28)])

}

#proportion_pos_slopes_per_subject_fight  

average_positive_slopes_across_subjects_fight <- mean(proportion_pos_slopes_per_subject_fight, na.rm = TRUE)
paste0 ("average_positive_slopes_across_subjects_fight = ", print(average_positive_slopes_across_subjects_fight))

# Note: a very small preponderance of positive slopes
```


# positive slopes when presented with a weapon and choosing to escape 
```{r}
proportion_pos_slopes_per_subject_escape <- 0
for (i in seq(nrow(positive_slopes_when_weapon_and_esccape))){
  
 proportion_pos_slopes_per_subject_escape[i] <- 
    
    sum(positive_slopes_when_weapon_and_esccape[i,c(2:28)]>0)/length(positive_slopes_when_weapon_and_esccape[i,c(2:28)])

}

#proportion_pos_slopes_per_subject_escape  

average_positive_slopes_across_subjects_escape <- mean(proportion_pos_slopes_per_subject_escape, na.rm = TRUE)
paste0 ("average_positive_slopes_across_subjects_escape = ", print(average_positive_slopes_across_subjects_escape))

# Note: a very small preponderance of positive slopes
```





# step to get a dataframe that denotes which trials amongst the ones where there is a negative contingency (i.e. non_weapon) had which CHOICE (fight or escape)
```{r}
slopes_non_weapon <- matrix(NA, nrow =  length(unique(Experiment3$subject)), ncol = length(unique(Experiment3$type2)))



for (j in seq(unique(Experiment3$subject))){
for (i in seq(unique(Experiment3$type2))) {
  
 # if(is.na(test[test$contingency=="weapon" & test$type2 == i ,]$anger_rating)==FALSE & #is.na(test[test$contingency=="non_weapon" & test$type2 == i ,]$anger_rating)==FALSE)
  
slopes_non_weapon[j,i] <- Experiment3[Experiment3$contingency=="non_weapon" & Experiment3$type2 == i & Experiment3$subject == j ,]$choice

#anger_slopes_df <- data.frame(cbind(subect = unique(Experiment3$subject), slopes))
 #colnames(anger_slopes_df) <- c("subject", unique(Experiment3$type2)) 

}
}

dim(slopes_non_weapon)
```


```{r}
positive_slopes_when_non_weapon_and_fight <- anger_slopes_df[slopes_non_weapon == "fight", ]
positive_slopes_when_non_weapon_and_escape <- anger_slopes_df[slopes_non_weapon == "escape", ]
dim(positive_slopes_when_non_weapon_and_fight )
dim(positive_slopes_when_non_weapon_and_escape)
```


# Actually, Song, this is a better way of presenting things here: 
If we take into acccount what happens in the non_weapon condition, the difference in positive slopes is fairly substantial: 
Given what I have done above
NUMBER OF POSITIVE SLOPES: 
a) Contingency WEAPON, Choice FIGHT: 
```{r}
nrow(anger_slopes_df[slopes_weapon == "fight", ])
```
b) Contingency WEAPON, Choice ESCAPE
```{r}
nrow(anger_slopes_df[slopes_weapon == "escape", ])
```

c) Contingency NON-WEAPON, Choice FIGHT: 
```{r}
nrow(anger_slopes_df[slopes_non_weapon == "fight", ])
```

d) Contingency NON-WEAPON, Choice FIGHT: 
```{r}
nrow(anger_slopes_df[slopes_non_weapon == "escape", ])
```



# as a table
```{r}
numbers_positive_slopes <- data.frame(rbind(nrow(anger_slopes_df[slopes_weapon == "fight", ]), nrow(anger_slopes_df[slopes_weapon == "escape", ]), nrow(anger_slopes_df[slopes_non_weapon == "fight", ]), nrow(anger_slopes_df[slopes_non_weapon == "escape", ])))
condition <- c("weapon_fight", "weapon_escape", "non_weapon_fight", "non_weapon_escape")
numbers_positive_slopes <- cbind( condition , numbers_positive_slopes)
colnames(numbers_positive_slopes) <- c("condition", "positive_slopes")
```

# as a graph
```{r}
numbers_positive_slopes %>% 
ggplot(aes(as.factor(condition), positive_slopes,count)) + geom_bar(stat="identity") +
  ylab ("trials with increase in anger (n)") +
  xlab  ("condition") +
  ggtitle  ("effect of affordance and choice on anger")
```
as a nicer graph
```{r}
library(stringr)

numbers_positive_slopes <- numbers_positive_slopes %>%
    mutate(chose_to_fight = str_detect(condition, "fight")) %>% 
  mutate(weapon = str_detect(condition, "non")) %>% 
  mutate(affordance =  ifelse(weapon=="TRUE", "non_weapon", "weapon"))

numbers_positive_slopes %>% 
ggplot(aes(affordance, positive_slopes,count, fill = chose_to_fight)) + 
  geom_bar(stat="identity") +
  ylab ("number of trials with increase in anger") +
  xlab  ("Affordance") +
  ggtitle  ("Anger increases when affordance and choice align")
```

# This does suggest a robust interaction effect. 


# What is the proportion of positive slopes per subject and across stimuli?
```{r}
proportion_pos_slopes_per_stimilus <- 0
for (i in seq(ncol(anger_slopes_df))){
  
  proportion_pos_slopes_per_stimilus[i] <- 
    
    sum(anger_slopes_df[,i]>0)/length(anger_slopes_df[, i])

}


paste0 ("The majority of stimuli lead to an increase in anger. Proportion_of_positive_slopes across_stimuli = ",sum(proportion_pos_slopes_per_stimilus[2:28]>0.5)/ncol(anger_slopes_df[,2:28]))



#paste0 ("average_positive_slopes_across_subjects = ", print(average_positive_slopes_across_subjects))

# Note: a very small preponderance of positive slopes
```









YOU CAN IGNORE THE BELOW, THESE ARE CHECKS THAT I HAVE DONE
```{r}

test <- Experiment3[Experiment3$subject == 1,]

test_slopes <- 0

for (i in seq(unique(test$type2))) {
  
 # if(is.na(test[test$contingency=="weapon" & test$type2 == i ,]$anger_rating)==FALSE & #is.na(test[test$contingency=="non_weapon" & test$type2 == i ,]$anger_rating)==FALSE)
  
 test_slopes[i] <-  test[test$contingency=="weapon" & test$type2 == i ,]$anger_rating - test[test$contingency=="non_weapon"& test$type2 == i,]$anger_rating
  
}

#test_slopes
#sum(test_slopes>0)/length(test_slopes)


```

some checks of the above
```{r}
check_for_loop <- list() #0
for (i in seq(unique(Experiment3$type2))) {
check_for_loop[i] <-test[test$contingency=="weapon" & test$type2 ==i, ]$anger_rating - test[test$contingency=="non_weapon"& test$type2 ==i,]$anger_rating == test_slopes[i]

}
#check_for_loop 
```

check the loop works by comparing this to the chunk below for subject 1
```{r}
sum(anger_slopes_df[1,c(2:28)]>0)/length(anger_slopes_df[1,c(2:28)]) == sum(test_slopes>0)/length(test_slopes) # this is from the chunk above for subject 1
```
