---
title: "Resting State Analyses"
author: "Argyris Stringaris"
date: "1/24/2022"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(ggplot2)
library(tidyverse)
library(Hmisc)
library(lme4)
```


Load Data
```{r}
Mmi.AllOpeningRestAndRandom_pymerInput.full <- read.csv("~/Downloads/Mmi-AllOpeningRestAndRandom_pymerInput-full.csv")

```


Mmi.AllOpeningRestAndRandom_pymerInput.full <- read.csv("~/Downloads/Mmi-AllOpeningRestAndRandom_pymerInput-full.csv")


```{r}
cutpoints<-c(0,.5,.9999)
table(cut_interval(1:100, 10))
test <- Mmi.AllOpeningRestAndRandom_pymerInput.full$Mood[1]  %>% 
  cut_interval()
Mmi.AllOpeningRestAndRandom_pymerInput.full$  

ids <- unique(Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject)    
initial_mood <-0  
for (i in seq(unique(Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject))){  
initial_mood[i] <- Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject == ids[i],]$Mood[1]  
}  
initial_mood

test <- cut_interval(initial_mood, 2)
```


1. get a sense of how different cohorts behave
Summary stats per cohort
```{r}

rest_sum_stats_age_depression <- Mmi.AllOpeningRestAndRandom_pymerInput.full %>% 
  group_by(Cohort) %>% 
  summarise(n_rest = n(), avg_age = mean(age), std_age = sd(age), avg_dep = mean(fracRiskScore), std_de = sd(fracRiskScore) )
rest_sum_stats_age_depression 


```

Plots of different cohorts
```{r}

annotation <- data.frame(
 # x = rep(c(0.6,0.9),length(rest_sum_stats$Cohort)),
 # y = c(400,500,length(rest_sum_stats$Cohort)),
  label = paste0("dep = ",round(rest_sum_stats_age_depression$avg_dep, 2), "+/-", round(rest_sum_stats_age_depression$std_de, 2)), 
  Cohort = rest_sum_stats_age_depression$Cohort, 
  size = 0.5
)  

# see distribution by depression
rest_distr_by_cohort <- Mmi.AllOpeningRestAndRandom_pymerInput.full %>% 
  ggplot(aes(x = Mood )) +
  ggtitle("mood distributions per cohort")+
  geom_histogram() +
  
  facet_wrap(Cohort ~ ., strip.position="top") +
 geom_text(data = annotation, aes( x = -Inf, y = 400, label = label,
                                   hjust   = -0.1,
                                   vjust   = -1,
                                   #color="orange", 
                                   
 ))
  
 
rest_distr_by_cohort

# Note: 1) NIMH sample seems to be the one with the lowest depression--we seem to have had a fair number of people participating  n = 116, 
#would have expected that many of them would have been depressed; 
#2) why is "recovery instructed" so flat?

```

2. get a sense of relationships between time and mood by group at the individual level
```{r}
Mmi.AllOpeningRestAndRandom_pymerInput.full$binary_dep <- ifelse(Mmi.AllOpeningRestAndRandom_pymerInput.full$fracRiskScore>=1.4375, 1, 0)

#cohort_1 <- Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Cohort == "COVID01",]


subsample_id <- sample(Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject, 30) # this allows you to look at different random subsamples each time

subsample_id <- Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject %in% subsample_id ,]

subsample_id  %>% ggplot(aes(x = Time,y = Mood, colour = binary_dep)) +
  geom_line() +
  facet_wrap(~Subject, ncol = 3)
```

 3. get a sense of relationships between time and mood by group at the cohort level
```{r}
rest_sum_stats_with_dep  <- Mmi.AllOpeningRestAndRandom_pymerInput.full %>% 
  group_by(cohort = Cohort, time = iRating, depression = as.factor(binary_dep)) %>% 
  summarise(n_rest = n(), avg_mood = mean(Mood), std_mood = sd(Mood), se_mood = std_mood/sqrt(n_rest), ci_mood = se_mood*qt(.975,n_rest-1)) 
rest_sum_stats_with_dep  

```

```{r}
mean_plot_per_cohort_by_dep <- ggplot(rest_sum_stats_with_dep , aes(x= time, y= avg_mood, colour = depression))+ 
  geom_line(aes(linetype = depression))+
  facet_wrap(~cohort)

mean_plot_per_cohort_by_dep 

mean_plot_per_cohort_by_dep + geom_ribbon(aes(x = time, y = avg_mood, ymin = avg_mood - ci_mood, ymax = avg_mood + ci_mood, xmin = min(time), xmax = max(time), 
                fill = "grey70", alpha = 0.3)) + ggtitle("Rest Mood per Cohort by Depression (=1, top 20%)")
```

4. get each person's FRS scores and each person's first and last mood to create slopes

Is the majority of slopes positive or negative? 
```{r}
ids <- unique(Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject)

unique_frac_risk_scores <- 0

mood_dif_per_subject <- 0

cohort_list <- 0

for (i in seq(unique(Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject))){
  
  unique_frac_risk_scores[i] <-   Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject == ids[i],]$fracRiskScore[1]
  
  
  mood_dif_per_subject[i] <- tail(Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject == ids[i],]$Mood, 1) -
          Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject == ids[i],]$Mood[1] 
  

  cohort_list[i] <-  Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Subject == ids[i],]$Cohort[1]
    
}


df_first_last_slopes <- data.frame(cbind(cohort_list, unique_frac_risk_scores, mood_dif_per_subject))

sum(mood_dif_per_subject<0)/length(mood_dif_per_subject) # Note: majority (68%) of slopes is negative
```


How do slopes correlate with depression risk?
```{r}
cor(as.numeric(unique_frac_risk_scores) , as.numeric(mood_dif_per_subject))

df_first_last_slopes %>% 
  
  ggplot(aes(x =  as.numeric(unique_frac_risk_scores), y =  as.numeric(mood_dif_per_subject))) +
  
  geom_point() +
  
  geom_smooth(method = "lm") +
  
  ylim(-1,1) +
  
  annotate("text", x = 2.8, y = 0.5, label = paste0("Pearson's r = ", cor(as.numeric(unique_frac_risk_scores) , as.numeric(mood_dif_per_subject))[1])) +   
  
  ggtitle("relationship between depression risk and slope (dif btw 1 and last mood rating")

```

Also doing the above broken down by depression
```{r}
df_first_last_slopes$binary_dep <- ifelse(df_first_last_slopes$unique_frac_risk_scores>=1.4375, "dep", "non_dep") # binarise depression


df_first_last_slopes %>% 
  
  ggplot(aes(x =  as.numeric(unique_frac_risk_scores), y =  as.numeric(mood_dif_per_subject), colour = as.factor(binary_dep))) +
  
  geom_point() +
  
  geom_smooth(method = "lm") +
  
  ylim(-1,1)
```



How do slopes relate to depression risk in a categorical analysis
```{r, error = TRUE}
df_first_last_slopes$what_slope <- ifelse(df_first_last_slopes$mood_dif_per_subject>0, "positive_slope", ifelse(df_first_last_slopes$mood_dif_per_subject==0,"zero_slope","negative_slope")) # binarise slopes


cross_tab_dep_slopes <-table(df_first_last_slopes$binary_dep,df_first_last_slopes$what_slope)

cross_tab_dep_slopes


prop_dep_slope <- data.frame(prop.table(cross_tab_dep_slopes,2)*100)

colnames(prop_dep_slope) <- c("depression", "slope", "percentage")

prop_dep_slope$slope <- factor(prop_dep_slope$slope)

slope_dep_chisq <- chisq.test(df_first_last_slopes$binary_dep,df_first_last_slopes$what_slope)
slope_dep_chisq_stats <- round(with(slope_dep_chisq, c(statistic, parameter, p.value)), 4)



```

plot the above
```{r, error = TRUE}
prop_dep_slope %>% 
  mutate(slope = fct_relevel(slope, #to show the progression of the slope direction
                            "negative_slope", "zero_slope", "positive_slope")) %>%
  ggplot(aes(x = slope, y = percentage, fill = depression)) +
  geom_bar(stat="identity", position=position_dodge()) +
  ggtitle("rest slope direction by depression level") +
  annotate("text", x = 3, y = 75, label = paste0("χ2 = ", slope_dep_chisq_stats[1], ", df = ", slope_dep_chisq_stats[2], ", p = ", slope_dep_chisq_stats[3] ))
```



check how percentage of neg slopes varies by cohort
```{r}
neg_slope_per_cohort <- df_first_last_slopes %>% 
  
  group_by(cohort_list) %>% 
  
  summarise(neg_slopes = sum(mood_dif_per_subject<0), tot_slopes = length(cohort_list), avg_mood_per_cohort = mean(as.numeric(unique_frac_risk_scores))) %>% 
  
   mutate(perc_neg_slope = (neg_slopes/tot_slopes)*100)
  
   


neg_slope_per_cohort 


neg_slope_per_cohort %>% 
  
  ggplot(aes(x = cohort_list, y = perc_neg_slope))+
  
  geom_bar(stat = "identity") +
  
  ggtitle("percentage negative slopes by cohort") +
  
  theme(axis.text.x = element_text(angle = 90)) +
  
  geom_hline(yintercept = 50 , linetype = "dashed", color = "red") 
```


6. LMEs
```{r}
# Model 1 with random intercept and slope for subject
mixed_rest_1 <- lmer (Mood~1+Time*(isMale+meanIRIOver20+totalWinnings+meanRPE+fracRiskScore+isAge0to16+isAge16to18+isAge40to100)+(Time|Subject), data = Mmi.AllOpeningRestAndRandom_pymerInput.full)
summary(mixed_rest_1)
```

```{r}
# Model 2 with random intercept and slope for subject and the same for cohort
mixed_rest_2 <- lmer (Mood~1+Time*(isMale+meanIRIOver20+totalWinnings+meanRPE+fracRiskScore+isAge0to16+isAge16to18+isAge40to100)+(Time|Subject)+(0+Time|Cohort), data = Mmi.AllOpeningRestAndRandom_pymerInput.full)
summary(mixed_rest_2)
```

```{r}

```

# Compare Models 1 & 2
```{r}
anova(mixed_rest_1, mixed_rest_2)
```



# Model 3 with random intercept and slope for subject nested within  cohort
```{r}

mixed_rest_3 <- lmer (Mood~1 + Time*(isMale+meanIRIOver20+totalWinnings+meanRPE+fracRiskScore+isAge0to16+isAge16to18+isAge40to100) + (Time|Cohort/Subject), data = Mmi.AllOpeningRestAndRandom_pymerInput.full)


```

# Compare Models 2 & 3
```{r}
anova(mixed_rest_2, mixed_rest_3)
```

# Model 4. Same as Model 1 without interaction
```{r}
mixed_rest_4 <- lmer (Mood~1 + fracRiskScore + Time*(isMale+meanIRIOver20+totalWinnings+meanRPE+isAge0to16+isAge16to18+isAge40to100)+(Time|Subject), data = Mmi.AllOpeningRestAndRandom_pymerInput.full)
summary(mixed_rest_4)


```

# Compare Models 1 & 4
```{r}
anova(mixed_rest_1, mixed_rest_4)
```



# Fit model across cohorts to test for interaction effect
```{r}
interaction_t_value <- 0#list()
Cohort <- unique(Mmi.AllOpeningRestAndRandom_pymerInput.full$Cohort)

for (i in seq(Cohort)){

interaction_t_value[i] <-  summary(lmer (Mood~1+Time*(isMale+meanIRIOver20+totalWinnings+meanRPE+fracRiskScore)+(Time|Subject), data = Mmi.AllOpeningRestAndRandom_pymerInput.full[Mmi.AllOpeningRestAndRandom_pymerInput.full$Cohort== Cohort[i],]))$coefficients[8,3]
 
}
t_values_interaction <- data.frame(cbind(Cohort, interaction_t_value))
t_values_interaction 

```


# 
```{r}

```


