---
title: "power_calculations_wellcome"
author: "Argyris Stringaris"
date: "01/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(faux)
library(simr)
library(lme4)
library(lmerTest)
library(stringr)
```


## Power calculations for Wellcome Application

Trying to simulate some of the models for the application. 
The basic form is Outcome = b1Group + b2Treatment + bxGroup1xTreatment + error
The idea is to 


```{r}
set.seed(1)
n_sims <- 1000 # we want 1000 simulations
p_vals <- c()
power_at_n <- c(0) # this vector will contain the power for each sample-size (it needs the initial 0 for the while-loop to work)
n <- 100 # sample-size and start at 100 as we can be pretty sure this will not suffice for such a small effect
n_increase <- 100 # by which stepsize should n be increased
i <- 2

power_crit <- .95
alpha <- .001

while(power_at_n[i-1] < power_crit){
  for(sim in 1:n_sims){
    group_TI <-  rnorm(n, 43, 15)
    group_VI <-  rnorm(n, 40, 15)
    group_TE <-  rnorm(n, 45, 15)
    group_VE <-  rnorm(n, 47, 15)
    
    participant <- c(1:(n*4))
    focus <- rep(c("internal", "external"), each = n*2)
    media <- rep(c("text", "visual"), each = n, times = 2)
    
    aov_dat <- data.frame(participant = participant, focus = focus, media = media, score = c(group_TI, group_VI, group_TE, group_VE))
    aov_dat$media_sum_num <- ifelse(aov_dat$media == "text", 1, -1) # apply sum-to-zero coding
    aov_dat$focus_sum_num <- ifelse(aov_dat$focus == "external", 1, -1) 
    lm_int <- lm(score ~ 1 + focus_sum_num + media_sum_num + focus_sum_num:media_sum_num, data = aov_dat) # fit the model with the interaction
    lm_null <- lm(score ~ 1 + focus_sum_num + media_sum_num, data = aov_dat) # fit the model without the interaction
    p_vals[sim] <- anova(lm_int, lm_null)$`Pr(>F)`[2] # put the p-values in a list
  }
    print(n)
    power_at_n[i] <- mean(p_vals < alpha) # check power (i.e. proportion of p-values that are smaller than alpha-level of .10)
    names(power_at_n)[i] <- n
    n <- n+n_increase # increase sample-size by 100 for low-resolution testing first
    i <- i+1 # increase index of the while-loop by 1 to save power and cohens d to vector
}

```


```{r}
power_at_n <- power_at_n[-1] 
```


```{r}
plot(as.numeric(names(power_at_n)), power_at_n, xlab = "Number of participants per group", ylab = "Power", ylim = c(0,1), axes = TRUE)
abline(h = .95, col = "red")
```

**Cohen's d to R2**
```{r}
#apparently d = 0.3 corresponds to R_sq = 0.022
M_A = 0
M_B = 50
SD_pooled = 15 # let's assume it is the same for both
d = 0.5

#d <- (M_A - M_B)/SD_pooled
MA <- d*SD_pooled + M_B
MA

```

**Let's create the dataset**
```{r}

    T_0_Cond_A <-  rnorm(n, 50, 15)
    T_0_Cond_B <-  rnorm(n, 50, 15)
    T_1_Cond_A <-  rnorm(n, 57.5, 15)
    T_1_Cond_B <-  rnorm(n, 50, 15)
    
    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*2)
    cond <- rep(c("A", "B"), each = n, times = 2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, score = c(T_0_Cond_A, T_0_Cond_B, T_1_Cond_A, T_1_Cond_B))
```

**check this works**
```{r}
df_for_sim %>% 
  group_by(cond, time) %>% 
  summarise(avg = mean(as.numeric(score), na.rm = TRUE))
```

**This is the bare bones of the simulation. You will need to do two more things. First, make the variables correlate within subject, i.e. the distributions for T0 and T1 must correlate within each condition. Then you will want to recover the R2 for which it may be best to extract the R2 of each model**
```{r}
n_sims <- 500
p_vals <- c()
n <- 200
for (i in 1:n_sims){

    T_0_Cond_A <-  rnorm(n, 50, 15)
    T_0_Cond_B <-  rnorm(n, 50, 15)
    T_1_Cond_A <-  rnorm(n, 57.5, 15)
    T_1_Cond_B <-  rnorm(n, 50, 15)
    
    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*2)
    cond <- rep(c("A", "B"), each = n, times = 2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, score = c(T_0_Cond_A, T_0_Cond_B, T_1_Cond_A, T_1_Cond_B))

  
  lm_int <- lm(score ~ 1 + cond + time + cond:time, data = df_for_sim) # fit the model with the interaction
    lm_null <- lm(score ~ 1 + cond + time, data = df_for_sim) # fit the model without the interaction
    p_vals[i] <- anova(lm_int, lm_null)$`Pr(>F)`[2] # put the p-values in a list
  
}
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```

create correlated variables. 
```{r}
library(faux)
    T_0_Cond_A <-  rnorm(n, 50, 15)
    T_0_Cond_B <-  rnorm(n, 50, 15)
    T_1_Cond_A <-  rnorm(n, 57.5, 15)
    T_1_Cond_B <-  rnorm(n, 50, 15)
    
    

```


```{r}
datA <- rnorm_multi(n = 200, 
                  mu = c(50, 57.5 ),
                  sd = c(15, 15),
                  r = c(0.5), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)
hist(datA$T_0_Cond_A)
colMeans(datA)
plot(datA$T_0_Cond_A, datA$T_1_Cond_A)
```




```{r}
datB <- rnorm_multi(n = 200, 
                  mu = c(50, 50.5 ),
                  sd = c(15, 15),
                  r = c(0.5), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)
hist(datB$T_0_Cond_B)
colMeans(datB)
plot(datB$T_0_Cond_B, datB$T_1_Cond_B)
```

```{r}
n = 200

    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*2)
    cond <- rep(c("A", "B"), each = n, times = 2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, score = c(datA$T_0_Cond_A, datB$T_0_Cond_B, datA$T_1_Cond_A, datB$T_1_Cond_B))
  
# check 
  a <- df_for_sim %>% filter(time == "time_0" & cond == "A")
  b <- df_for_sim %>% filter(time == "time_1" & cond == "A")
  cor(a$score,b$score)
  df_for_sim %>% group_by(time, cond) %>% summarise (avg = mean(score))
```

**this is more appropriate than the above as I allow the measures to correlate within person**
```{r}
n_sims <- 1000
p_vals <- c()
n <- 120

for (i in 1:n_sims){

    datA <- rnorm_multi(n = n, # create a two correlated variables
                  mu = c(50, 57.5 ),
                  sd = c(15, 15),
                  r = c(0.5), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

    datB <- rnorm_multi(n = n,  # create a two correlated variables
                  mu = c(50, 50.5 ),
                  sd = c(15, 15),
                  r = c(0.5), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)

    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*2)
    cond <- rep(c("A", "B"), each = n, times = 2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, surprise = c(datA$T_0_Cond_A, datB$T_0_Cond_B, datA$T_1_Cond_A, datB$T_1_Cond_B))
  
   lm_int <- lm(surprise ~ 1 + cond + time + cond:time, data = df_for_sim) # fit the model with the interaction
    lm_null <- lm(surprise ~ 1 + cond + time, data = df_for_sim) # fit the model without the interaction
    p_vals[i] <- anova(lm_int, lm_null)$`Pr(>F)`[2] # put the p-values in a list
  
}
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```


```{r}
n_sims <- 1000
p_vals <- c()
n <- 100
r_for_scores = 0.5
#r_for_surprise = 0.3


for (i in 1:n_sims){

    datA_high <- rnorm_multi(n = n, # create a two correlated variables
                  mu = c(50, 57.5 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

    datA_low <- rnorm_multi(n = n, # create a two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)
 
    datB_high <- rnorm_multi(n = n,  # create a two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)
    
    datB_low <- rnorm_multi(n = n,  # create a two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)

    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*4)
    cond <- rep(c("A", "B"), each = n, times = 4)
    surprise <- rep(c("high", "low"), each = n*2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, surprise = surprise,
  score = c(datA_high$T_0_Cond_A, 
            datA_low$T_0_Cond_A,
            datB_high$T_0_Cond_B, 
            datB_low$T_0_Cond_B,
                                                                                           
            datA_high$T_1_Cond_A, 
            datA_low$T_1_Cond_A,
            datB_high$T_1_Cond_B, 
            datB_low$T_1_Cond_B))
  
   lm_int <- lm(score ~ 1 + cond*time*surprise, data = df_for_sim) # fit the model with the interaction
    lm_null <- lm(score ~ 1 + cond*time , data = df_for_sim) # fit the model without the interaction
    p_vals[i] <- anova(lm_int, lm_null)$`Pr(>F)`[2] # put the p-values in a list
  
}
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```

```{r}
df_for_sim %>% group_by(time, cond, surprise) %>% summarise(avg= (mean(score)))
```

# try to see what you need in order to recover a diffrence of d = 0.5 for the interaction
```{r}
n_sims <- 1000
p_vals <- c()
n <-180
r_for_scores = 0.5
#r_for_surprise = 0.3


for (i in 1:n_sims){

    datA_high <- rnorm_multi(n = n, # create two correlated variables
                  mu = c(50, 57.5 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

    datA_low <- rnorm_multi(n = n, # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)
 
    datB_high <- rnorm_multi(n = n,  # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)
    
    datB_low <- rnorm_multi(n = n,  # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)

    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*4)
    cond <- rep(c("A", "B"), each = n, times = 4)
    surprise <- rep(c("high", "low"), each = n*2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, surprise = surprise,
  score = c(datA_high$T_0_Cond_A, 
            datA_low$T_0_Cond_A,
            datB_high$T_0_Cond_B, 
            datB_low$T_0_Cond_B,
                                                                                           
            datA_high$T_1_Cond_A, 
            datA_low$T_1_Cond_A,
            datB_high$T_1_Cond_B, 
            datB_low$T_1_Cond_B))
  
   lm_int <- lm(score ~ 1 + cond*time + time*surprise, data = df_for_sim) # fit the model with the interaction
   # lm_null <- lm(score ~ 1 + cond*time , data = df_for_sim) # fit the model without the interaction
    p_vals[i] <- summary(lm_int)$coefficients[length(lm_int$coefficients),4] # put the p-values in a list
  
}
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```


```{r}
model_1 <- lmer(score~time*cond + time*surprise + (1|subject), data = df_for_sim )
summary(model_1)$coef
```

```{r}
powerSim(model_1, fixed("timetime_1:surpriselow", "z"), nsim=50)
```


```{r}
df_for_sim %>% group_by(time, cond, surprise) %>% summarise(avg= (mean(score)))
```

# try to see what you need in order to recover a diffrence of d = 0.5 for the interaction
```{r}
n_sims <- 1000
p_vals <- c()
n <-200
r_for_scores = 0.5
#r_for_surprise = 0.3


for (i in 1:n_sims){

    datA_high <- rnorm_multi(n = n, # create two correlated variables
                  mu = c(50, 57.5 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

    datA_low <- rnorm_multi(n = n, # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)
 
    datB_high <- rnorm_multi(n = n,  # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)
    
    datB_low <- rnorm_multi(n = n,  # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)

    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*4)
    cond <- rep(c("A", "B"), each = n, times = 4)
    surprise <- rep(c("high", "low"), each = n*2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, surprise = surprise,
  score = c(datA_high$T_0_Cond_A, 
            datA_low$T_0_Cond_A,
            datB_high$T_0_Cond_B, 
            datB_low$T_0_Cond_B,
                                                                                           
            datA_high$T_1_Cond_A, 
            datA_low$T_1_Cond_A,
            datB_high$T_1_Cond_B, 
            datB_low$T_1_Cond_B))
  
   lmer_int <- lmer(score~time*cond + time*surprise + (1|subject), data = df_for_sim ) # fit the model with the interaction
   # lm_null <- lm(score ~ 1 + cond*time , data = df_for_sim) # fit the model without the interaction
    p_vals[i] <- summary(lmer_int)$coef[6,5] # put the p-values in a list
  
}  
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```





```{r}
n_sims <- 1000
p_vals <- c()
n <-200
r_for_scores = 0.5
#r_for_surprise = 0.3


for (i in 1:n_sims){

    datA_high <- rnorm_multi(n = n, # create two correlated variables
                  mu = c(50, 57.5 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

    datA_low <- rnorm_multi(n = n, # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)
 
    datB_high <- rnorm_multi(n = n,  # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)
    
    datB_low <- rnorm_multi(n = n,  # create two correlated variables
                  mu = c(50, 50 ),
                  sd = c(15, 15),
                  r = c(r_for_scores), 
                  varnames = c("T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE)

    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*4)
    cond <- rep(c("A", "B"), each = n, times = 4)
    surprise <- rep(c("high", "low"), each = n*2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, surprise = surprise,
  score = c(datA_high$T_0_Cond_A, 
            datA_low$T_0_Cond_A,
            datB_high$T_0_Cond_B, 
            datB_low$T_0_Cond_B,
                                                                                           
            datA_high$T_1_Cond_A, 
            datA_low$T_1_Cond_A,
            datB_high$T_1_Cond_B, 
            datB_low$T_1_Cond_B))
  
   lmer_int <- lmer(score~time*cond + time*surprise + (1|subject), data = df_for_sim ) # fit the model with the interaction
   # lm_null <- lm(score ~ 1 + cond*time , data = df_for_sim) # fit the model without the interaction
    p_vals[i] <- summary(lmer_int)$coef[6,5] # put the p-values in a list
  
}  
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```





```{r}

avg_A <- 50
avg_B <- 56
n <-100
r_for_scores = 0.5

datX <- rnorm_multi(n = n, # create many correlated variables
                  mu = c(avg_A, avg_B , avg_A, avg_B , avg_A, avg_B , avg_A, avg_B , avg_A, avg_B , avg_A, avg_B),#c(50, 55 , 50, 55 , 50, 55 , 50, 55 , 50, 55 ,50, 55), #c(avg_A, avg_B , avg_A, avg_B , avg_A, avg_B , avg_A, avg_B , avg_A, avg_B , avg_A, avg_B),
                  sd = c(15, 15,15, 15,15, 15,15, 15,15, 15,15, 15),
                  r = c(r_for_scores), 
                  varnames = c("A", "B", "A", "B", "A", "B", "A", "B", "A", "B", "A", "B"),
                  empirical = FALSE)

#datX <- data.frame(cbind(id), datX)
```


```{r}
datX_long <- datX %>%
  pivot_longer(
    cols = starts_with(c("A", "B")),
    names_to = "cond", 
    #names_prefix = "Condd",
    values_to = "score")
id <- rep(1:n, each=12)
time<- rep(1:12, n)
datX_long <- data.frame(cbind(id,time, datX_long))




# test <- str_replace(datX_long$cond , "Cond_A_1", "A")
# test <- str_replace(test , "Cond_A_2", "A")
# test <- str_replace(test , "Cond_A_3", "A")
# test <- str_replace(test , "Cond_A_4", "A")
# test <- str_replace(test , "Cond_A_5", "A")
# test <- str_replace(test , "Cond_A_6", "A")
# test <- str_replace(test , "Cond_B_1", "B")
# test <- str_replace(test , "Cond_B_2", "B")
# test <- str_replace(test , "Cond_B_3", "B")
# test <- str_replace(test , "Cond_B_4", "B")
# test <- str_replace(test , "Cond_B_5", "B")
# test <- str_replace(test , "Cond_B_6", "B")
# test[1:10]

# datX_long$cond <- test

datX_long %>% 
  group_by(cond,time) %>% 
  summarise(avg= mean(score), std = sd(score))

```



```{r}
mod_1 <- lmer(score~time*cond  + (1|id), data = datX_long)
summary(mod_1)$coef
```

```{r}
powerSim(mod_1, fixed("condB", "z"), nsim=50)
```







```{r}
gm1 <- glmer(cbind(incidence, size - incidence) ~ period + (1 | herd) + (1|obs), data=cbpp,
    family=binomial)
summary(gm1)$coef
```




```{r}
dat_test_A <- rnorm_multi(n = n, 
                  mu = c(30, 50, 57.5),
                  sd = c(10, 15, 15),
                  r = c(0.3, 0.3, 0.5), 
                  varnames = c("surprise", "T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

cor(dat_test_A)
```

```{r}
dat_test_B <- rnorm_multi(n = n, 
                  mu = c(30, 50, 50),
                  sd = c(10, 15, 15),
                  r = c(0.1, 0.1, 0.5), 
                  varnames = c("surprise", "T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

cor(dat_test_B)

```



```{r}
n_sims <- 1000
p_vals <- c()
n <- 500
r_for_scores = 0.8
r_for_surprise = 0.8


for (i in 1:n_sims){

dat_test_A <- rnorm_multi(n = n, 
                  mu = c(30, 50, 57.5),
                  sd = c(10, 15, 15),
                  r = c(0.3, 0.3, 0.5), 
                  varnames = c("surprise", "T_0_Cond_A", "T_1_Cond_A"),
                  empirical = FALSE)

dat_test_B <- rnorm_multi(n = n, 
                  mu = c(30, 50, 50),
                  sd = c(10, 15, 15),
                  r = c(0.1, 0.1, 0.5), 
                  varnames = c("surprise", "T_0_Cond_B", "T_1_Cond_B"),
                  empirical = FALSE) 


    subject <- c(1:(n*4))
    time <- rep(c("time_0", "time_1"), each = n*2)
    cond <- rep(c("A", "B"), each = n, times = 2)
    
  df_for_sim <-  data.frame(subject = subject, time = time, cond = cond, score = c(dat_test_A$T_0_Cond_A, dat_test_B$T_0_Cond_B, dat_test_A$T_1_Cond_A, dat_test_B$T_1_Cond_B),
                            surprise = dat_test_A$surprise)
  
   lm_int <- lm(score ~ 1 + time*surprise + time*cond, data = df_for_sim[cond=="A",]) # fit the model with the interaction
    lm_null <- lm(score ~ 1 + time*cond, data = df_for_sim[cond=="A",]) # fit the model without the interaction
    p_vals[i] <- anova(lm_int, lm_null)$`Pr(>F)`[2] # put the p-values in a list
  
}
sum(p_vals<0.05)/length(p_vals)
length(p_vals)
```




```{r}
n_sims <- 1000 # we want 1000 simulations
p_vals <- c()
  for(sim in 1:n_sims){
    group_TI <-  rnorm(n, 43, 15)
    group_VI <-  rnorm(n, 40, 15)
    group_TE <-  rnorm(n, 45, 15)
    group_VE <-  rnorm(n, 47, 15)
    
    participant <- c(1:(n*4))
    focus <- rep(c("internal", "external"), each = n*2)
    media <- rep(c("text", "visual"), each = n, times = 2)
    
    aov_dat <- data.frame(participant = participant, focus = focus, media = media, score = c(group_TI, group_VI, group_TE, group_VE))
    aov_dat$media_sum_num <- ifelse(aov_dat$media == "text", 1, -1) # apply sum-to-zero coding
   aov_dat$focus_sum_num <- ifelse(aov_dat$focus == "external", 1, -1) 
    lm_int <- lm(score ~ 1 + focus_sum_num + media_sum_num + focus_sum_num:media_sum_num, data = aov_dat) # fit the model with the interaction
    lm_null <- lm(score ~ 1 + focus_sum_num + media_sum_num, data = aov_dat) # fit the model without the interaction
    p_vals[sim] <- anova(lm_int, lm_null)$`Pr(>F)`[2] # put the p-values in a list
  }
p_vals
```



```{r}
n = 100
n_sims = 500
est_null <- 0
est_two_var <- 0
est_inter <- 0
for(i in 1:n_sims){
test <- rnorm_multi(n = n , mu = 100, sd = 15, r = 0.5, varnames = letters[20:22])
est_null[i] <- summary(lm(test[,1]~test[,2] + test[,3], test))$coefficients[2,4]
est_two_var[i] <- summary(lm(test[,1]~test[,2] + test[,3], test))$coefficients[3,4]
est_inter[i] <- summary(lm(test[,1]~test[,2] + test[,3] + test[,2]*test[,3], test))$coefficients[4,4]
estimates_df <- data.frame(est_null, est_two_var, est_inter)
}

colMeans(estimates_df)
```
colmeans




```{r}
x <- rnorm(100, 100, 15)
#y <- rnorm(100, 100, 15)
e <- rnorm(100, 0, 1)

for(i in 1:n_sims){
y <- 1 +  0.5*x + e

test <- data.frame(cbind(y,x))
estimate[i] <-summary(lm(y~x))$coefficients[2,1]

}
mean(estimate)

```

```{r}
install.packages("simr")
library(simr)
cbpp$obs <- 1:nrow(cbpp)
gm1 <- glmer(cbind(incidence, size - incidence) ~ period + (1 | herd) + (1|obs), data=cbpp,
    family=binomial)
summary(gm1)$coef
doTest(gm1, fixed("period", "lr"))
doTest(gm1, fixed("period2", "z"))
gm2 <- glmer(cbind(incidence, size - incidence) ~ period + size + (1 | herd), data=cbpp,
    family=binomial)
doTest(gm2, fixed("size", "z"))
fixef(gm2)["size"] <- 0.05
powerSim(gm2, fixed("size", "z"), nsim=50)

fm1 <- lmer(angle ~ recipe:temp + (1|recipe:replicate), data=cake, REML=FALSE)
doTest(fm1, fixed("recipe", "lr"))
fixef(fm1)["recipeA:temp"] <- 0.5
powerSim(fm1, fixed("recipeA:temp", "z"), nsim=50)

data(budworm, package="pbkrtest")
bw1 <- glmer(cbind(ndead, ntotal-ndead) ~ dose*sex, family="binomial", data=budworm)
summary(bw1)$coef
doTest(bw1, compare(. ~ dose + sex))
doTest(bw1, fixed("dose:sexmale", "z"))
fixef(bw1)["dose:sexmale"] <- 0.5

```

```{r}
subj <- factor(1:10)
class_id <- letters[1:5]
```


```{r}
time <- 0:2
group <- c("control", "intervention")

subj_full <- rep(subj, 15)
class_full <- rep(rep(class_id, each=10), 3)
time_full <- rep(time, each=50)
group_full <- rep(rep(group, each=5), 15)

covars <- data.frame(id=subj_full, class=class_full, treat=group_full, time=factor(time_full))

covars

tail(covars)
str(covars$id)
```

```{r}
fixed <- c(5, 0, 0.1, 0.2, 1, 0.9)
rand <- list(0.5, 0.1)

## residual variance
res <- 2
```

```{r}
model <- makeLmer(y ~ treat*time + (1|class/id), fixef=fixed, VarCorr=rand, sigma=res, data=covars)
model
```

```{r}

```
```{r}
fm2 <- lmer(angle ~ recipe + poly(temp, 2) + (1|recipe:replicate), data=cake, REML=FALSE)
summary(fm2)$coef
```

```{r}
fixef(fm2)
```

